\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2015-09-21 methodology.tex
more edits

The ExtMgr runs internally a periodic \emph{EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period of time, as well as cleaning up the corresponding extent records in \texttt{ExtentCenter}. In addition, the ExtMgr runs a periodic \emph{extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the ENs. The \psharp test harness mocks the original network engine in vNext by overriding its interface. This enables the mocked network engine to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.

@@

Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it running in a real distributed environment and communicating with real ENs.

@@

System correctness should \emph{not} hinge on the frequency of any individual timer. Instead, all nondeterminism due to timing-related events must be delegated to \psharp. To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}. Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.

@@

Figure~\ref{fig:testing_driver} illustrates how \psharp can be used for injecting nondeterministic failures. The \texttt{TestingDriver} machine can invoke the \psharp method \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode}machines. Let $n_f$ be this chosen machine. \texttt{TestingDriver} can emulate a failure by sending $n_f$ a \texttt{FailEvent} event. The $n_f$ machine handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}). \texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for unit testing, which maximizes code re-use. Next, $n_f$ notifies the liveness monitor that it has failed. Finally, $n_f$ calls the \psharp method \texttt{Halt()}, which causes $n_f$ to terminate. \psharp is able to not only enumerate interleavings of asynchronous event handlers, but also calls to \texttt{Nondet()}, and thus failure injections.

@@

However, the original \psharp trace did not include sufficient detail to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran again the test harness. No bugs were found during 100,000 iterations, a process that required only tens of minutes.

@@

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from \texttt{ExtentNodeMap}; (iii) \texttt{ExtentCenter} is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy. In production, such a bug can cause a very serious incident of customer data loss.

The culprit is in (iv), where ExtMgr receives a sync report from EN$_0$ after deleting the EN. This may occur in \psharp due to arbitrary interleaving of events. It may also occur, albeit much less frequently, during stress testing due to messages being delayed in the network. This explains why the bug only occurs from time to time during stress testing and requires long executions to manifest. In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit to be identified promptly, and the fix to be tested effectively and thoroughly, all of which have the potential to vastly increase the productivity of distributed storage system development.

>>>>

ExtMgr internally runs a periodic \emph{EN expiration loop} that is responsible for removing ENs that have been missing heartbeats for an extended period of time, as well as cleaning up the corresponding extent records in \texttt{ExtentCenter}. In addition, ExtMgr runs a periodic \emph{extent repair loop} that examines all the \texttt{ExtentCenter} records, identifies extents with missing replicas, schedules extent repair tasks and sends them to the ENs.

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the ENs. The \psharp test harness mocks the original network engine in vNext by overriding its interface. This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.

@@

Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.

@@

System correctness should \emph{not} hinge on the frequency of any individual timer. Instead, all nondeterminism due to timing-related events must be delegated to \psharp. To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach also used in previous work~\cite{desai2015building}. Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.

@@

Figure~\ref{fig:testing_driver} illustrates how \psharp can be used for injecting nondeterministic failures. The \texttt{TestingDriver} machine uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode} machines. If \texttt{n} is the chosen machine, \texttt{TestingDriver} emulates a failure by sending a \texttt{FailEvent} event to \texttt{n}. Machine \texttt{n} handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}). \texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for simulating failures in its unit tests; we simply re-use this functionality. Next, \texttt{n} notifies the liveness monitor that it has failed. Finally, machine \texttt{n} calls the \psharp method \texttt{Halt()}, which causes \texttt{n} to terminate. \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.

@@

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debuggin outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found for 100,000 executions, a process that only took on the order of tens of minutes.

@@

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from \texttt{ExtentNodeMap}; (iii) \texttt{ExtentCenter} is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr would still think that all replicas are healthy. In production, such a bug can cause a very serious incident of customer data loss.

The culprit is in step (iv), where ExtMgr receives a sync report from EN$_0$ after deleting the EN. This interleaving is exposed quickly by \psharp's testing engine that has the control to arbitrarily interleave events. It may also occur, albeit much less frequently, during stress testing due to messages being delayed in the network. This explains why the bug only occurs from time to time during stress testing and requires long executions to manifest. In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit to be identified promptly, and the fix to be tested effectively and thoroughly, all of which have the potential to vastly increase the productivity of distributed storage system development.
